### 2017-02-15 pedestrian_forecasting_iros_2017.tex
Here's the paper repo

\documentclass[letterpaper,10pt,conference]{ieeeconf}
%\usepackage{times}

\IEEEoverridecommandlockouts 
\overrideIEEEmargins   % Needed to meet printer requirements.
\usepackage[T1]{fontenc}

%  For math environment 
\usepackage{amsmath,amssymb}
%\usepackage[linesnumbered]{algorithm2e}
\usepackage{algpseudocode}
\usepackage{algorithm}
%\usepackage{algorithm2e}
\usepackage{caption}
%  For todo notes
\usepackage{todonotes}
%  For drawing Bayesian networks
\usepackage{tikz}
\usetikzlibrary{arrows,positioning}
\usepackage{tikz-cd}
\usepackage[justification=justified,font=footnotesize]{caption}
\usepackage{subcaption}
\usepackage{graphicx}
\usepackage{array}
\newcolumntype{M}[1]{>{\centering\arraybackslash}m{#1}}
\usepackage{comment}
\usepackage{color}  % For Highlighting

%  For theorems, etc
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}

% numbers option provides compact numerical references in the text. 
%\usepackage[numbers]{natbib}
%\usepackage{multicol}
\usepackage[bookmarks=true]{hyperref}
\newcommand{\Ram}[1]{\textcolor{red}{#1}}
\newcommand{\Hoj}[1]{\textcolor{blue}{#1}}

% Custom commands
\newcommand{\pder}[2]{\ensuremath{\frac{\partial #1}{\partial #2}}}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}

\begin{document}

% paper title
\title{Real-Time Certified Probabilistic Pedestrian Forecasting}

%%% first author
\author{Henry O. Jacobs, Owen Hughes, Matthew Johnson-Roberson, and Ram Vasudevan
 \thanks{ H.O. Jacobs, O. Hughes, and R.~Vasudevan are with the Department of Mechanical Engineering, University of Michigan, Ann Arbor, MI 48109
{\scriptsize \{\texttt{hojacobs,owhughes,ramv}\}~\texttt{@umich.edu}}}
\thanks{M. Johnson-Roberson is with the Department of Naval and Marine Engineering, University of Michigan, Ann Arbor, MI 48109\newline
{\scriptsize \texttt{mattjr@umich.edu}}}
}

% You will get a Paper-ID when submitting a pdf file to the conference system
%\newcommand{\umich}{ \begin{affiliation} Dept. of Mechanical Eng. \\ University of Michigan \\ Ann Arbor, Michigan \end{affiliation}}
%\author{
%	Henry O. Jacobs \\
%	\umich \\
%	\email{hojacobs@umich.edu}
%	\and
% 	Owen Hughes \\
%	\umich \\
%	\email{owhughes@umich.edu}
% 	\and Ram Vasudevan \\
%	\umich \\
%	\email{ramv@umich.edu} \\
%}

\maketitle

\begin{abstract}
That autonomous vehicles will come to dominate our streets is imminent.
This motivates the need for a real-time probabilistic forecasting algorithm for pedestrians, cyclists, and other agents,
as it forms a necessary step in assessing the risk (and therefore the cost) we should expect to incur.
In this paper, we present a novel approach to probabilistic forecasting for pedestrians based on weighted sums of ordinary differential equations.
The resulting algorithm is embarrassingly parallel, and trained on historical trajectory information within a fixed scene.
When compared with MDP-based methods, our algorithm appears to be superior from the standpoint of precision and recall.
\end{abstract}

%\IEEEpeerreviewmaketitle

\section{Introduction}

Autonomous systems are increasingly being deployed in and around humans. 
The ability to accurately model and anticipate human behavior is critical to maximizing safety, confidence, and effectiveness of these systems in human-centric environments. 
The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of an action over a set of possible behaviors. 
Since the set of plausible human behaviors is vast, this work focuses on anticipating the possible future locations of people within a bounded area. 
This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).

With a particular focus on the AV application several additional design criteria become important. 
First, false negative rates for unoccupied regions must be minimized. 
The misclassification of space in this way has obvious safety issues. 
Second, speed is paramount. 
To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories. 
The margins between humans and vehicles are small and both frequently operate at the boundary. 
Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.
This paper presents an algorithm for real-time, long-term prediction of pedestrian behavior which can subsequently be used by autonomous agents.
As depicted in Figure \ref{fig:gates-1-2}, this method works quickly to generate predictions that are precise while reducing the likelihood of false negative detections.

\begin{figure}
	\centering
	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=55.png}
		\caption{$t=1.83s$}
	\end{subfigure}
		\begin{subfigure}[b]{.45\linewidth}
			\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=145.png}
			\caption{$t=4.83$}
		\end{subfigure}

	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=235.png}
		\caption{$t=7.83s$}
	\end{subfigure}
	\begin{subfigure}[b]{.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/FirstPage/gates_1_2_t=345.png}
		\caption{$t=11.5s$}
	\end{subfigure}
	\caption{The performance of the presented algorithm captures the most probable routes that a pedestrian chooses. In this example which was captured at thirty frames per second, the presented algorithm took 0.0158s per frame in a Python implementation.  In this figure, the dot is the start point of the test trajectory, the diamond is the position at time $t$, and the X is the end of the trajectory.}
	\label{fig:gates-1-2}
\end{figure}

Autonomous systems are increasingly being deployed in and around humans.
 The ability to accurately model and anticipate human behavior is of great value to maximize safety, confidence, and effectiveness in human-centric environments. 
 The stochasticity of humans necessitates a probabilistic approach to capture the likelihood of possible action over a set a of possible behaviors. 
 Since the set of plausible behaviors in arbitrary situations may be vast, this work constrains the problem to anticipating the possible future locations of people within a bounded area. This problem is critical in many application domains such as enabling personal robots to navigate in crowded environments, managing pedestrian flow in smart cities, and synthesizing safe controllers for autonomous vehicles (AV).

 With a particular focus on the AV application several additional design criteria become important. 
 First, false negative rates for unoccupied regions must be minimized. 
 The misclassification of space in this way has obvious safety issues. 
 Second, speed is paramount. 
 To effectively use human prediction within a vehicle control loop prediction rates must be commensurate with the speed at which humans change trajectories. 
 The margins between humans and vehicles are small and both frequently operate at the boundary. 
 Finally, long-time horizon forecasting is preferable since this improves robot predictability, reduces operation close to safety margins, prevents the need for overly conservative or aggressive controllers and makes high-level goal planning more feasible.

This paper presents an algorithm for real-time, long-term prediction of pedestrian behavior which can subsequently be used by autonomous agents.
As depicted in Figure \ref{fig:}, this approach relies upon ...

%First, since safety is paramount, generated predictions must avoid misclassifying areas as unoccupied even though humans are present. 
%Though this may unduly reduce the precision of the algorithm

\subsection{Background}
Most forecasting algorithm are well characterized by the underlying evolution model they adopt.
Such models come in a variety flavors, and are adapted to the task at hand (e.g. crowd modeling \cite{Helbing1992}).
This paper is focused on the construction of useful motion models for pedestrians that can aide the task of real-time forecasting for autonomous agents.
The simplest approach to forecasting with motion models forward integrates a Kalman filter \cite{kalman1960new} based upon the observed heading.
Over short time scales this method may perform well, but the resulting distribution devolves into an imprecise Gaussian mass over longer time scales.
This is a typical property of the extended Kalman filter and the Kalman filter with deterministic nonlinear drift \cite{Schneider2013}.
In particular, such models are less useful for forecasts beyond two seconds, especially when a pedestrian turns.
Nonetheless, these stochastic linear models serve as a reasonable default in the absence of any contextual knowledge.
For example, \cite{Kooji2014} uses a motion model of this variety, parametrized by latent variables, such as the pedestrian's awareness of nearby vehicles, to predict when and how a pedestrian may cross a street without leveraging any environmental context.

More sophisticated models that attempt to leverage environmental data include inverse optimal control (IOC) based models \cite{Ziebart2008,Ziebart2009,Kitani2012,Xie2013,Karasev2016}.
These IOC models have the desirable property of attributing intention and goal-seeking behavior to the agents.
For example, \cite{Kitani2012} extracts a Markov Decision Process (MDP) evolving on a finite 2D lattice by training on a small collection of features and trajectories.
The resulting MDP is light-weight since it is parametrized by only a small collection of coefficients equal in number to that of the feature maps.
Moreover, given an initial and final state, the probability distribution of the pedestrian at intermediate states is computable using matrix multiplication.
The speed of this computation, makes the algorithm of \cite{Kitani2012} a reasonable baseline for comparison for the algorithm that is presented in this paper.

The approach presented in \cite{Kitani2012} has been generalized in a variety of ways.
For example, time-dependent information, such as traffic signals, are incorporated in \cite{Karasev2016}, by relaxing the Markov property and considering a switched Markov process.
Other generalizations in \cite{Karasev2016} include replacing the finite-state space with a continuous one, and using a Markov jump process in the motion model.
Unfortunately the desired posteriors are difficult to compute in closed form, and as a result a sampling based method: a Rao-Blackwellized particle filter~\cite{Doucet2000} is employed.
The resulting accuracy of such methods can only be known in a probabilistic sense in that the error bounds are themselves random variables.
Moreover, accuracy can come at a large computational expensive which is prohibitive during real-time control.

One limitation of IOC models occurs in cases where there are many locally optimal solutions between a start and end goal.
In these cases, IOC type methods can yield non-robust and imprecise behavior.
This can occur, for example, when agents make sharp turns due to intermediate criteria on the way toward reaching their final destination.
To address this, \cite{Ballan2016} adopt an empiricists approach, computing ``turning maps'' and attempting to infer how agents behave in a given patch based on its feature and the behavior of previous agents on similar patches.
The motion model is a Markov jump process and the relevant posteriors were approximated using sample based techniques similar to \cite{Karasev2016}.
The objective of \cite{Ballan2016} is not only prediction, but the development of a motion model learned on one scene that could then subsequently be transferred to other scenes. 
This requires representations of ``objects'' in the scene that do not depend rigidly on the finitely many labels an engineer managed to think of in a late-night brainstorming session.
For example, such an algorithm should be able to infer how to behave near a roundabout by using prior knowledge of more fundamental building blocks like pavement, curbs, and asphalt.

Along similar lines to \cite{Ballan2016}, \cite{Walker2014} constructed an unsupervised approach towards forecasting.
As before, the motion model was a Markov jump process, and the training set was a collection unlabeled videos.
Unlike all the approaches mentioned thus far, the agents in \cite{Walker2014} were not manually specified.
They were learned by detecting which sort of patches of video were likely to move, and how.
The resulting predictions outperformed \cite{Kitani2012} when comparing the most likely path with the ground truth using the mean Hausdorff distance.
As in all methods mentioned thus far, computational speed and accuracy of any predicted posteriors were not a concern of \cite{Walker2014}, so no such results were reported.
However, since the motion model was a Markov jump process which required the application of a sample based technique, we should expect the same painful trade-off between error and speed to occur as in \cite{Karasev2016} and \cite{Ballan2016}.

\subsection{Contributions}

The primary contributions of this paper are:
\begin{itemize}
	\item An accurate motion model for pedestrian forecasting.
	\item An expedient method for computing approximations of relevant posteriors generated by our motion model.
	\item Hard error bounds on the proposed approximations.
\end{itemize}

For clarification, we should mention that there are a number of things that we do not do.
For example, we do not concern ourselves with detection and tracking. 
Nor do we concern ourselves with updating our prediction as new data comes along \cite{Kitani2012}.
We largely work in a 2D environment with a bird's eye view, operating under the assumption that the data has been appropriately transformed by a third party.
This is in contrast to \cite{Karasev2016}, which operates from first person video.
While it would be a straight forward extension to consider such things, it would detract from the presentation.
%Such questions are important, but beyond the scope of this paper, which concerns the underlying dynamic model.

The rest of the paper is organized as follows:
\begin{itemize}
	\item \S \ref{sec:model} describes our motion model as a Bayesian network.
	\item \S \ref{sec:efficient} describes how to compute probability densities for an agent's position efficiently.
	\item \S \ref{sec:implementation} demonstrates the model by training and testing it on the Stanford drone dataset \cite{Robicquet2016}.
\end{itemize}

\section{Model}\label{sec:model}
This paper's goal is to generate a time-dependent probability density over $\mathbb{R}^2$, which predicts the true location of an agent in the future.
The input to the algorithm at runtime is a noisy measurement of position and velocity, $\hat{x}_0, \hat{v}_0 \in \mathbb{R}^2$.
If the (unknown) location of agent at time $t$ is given by $x_t \in \mathbb{R}^2$, then the distribution we seek is the posterior $\rho_t(x_t) := \Pr( x_t \mid \hat{x}_0, \hat{v}_0 )$.

To numerically compute $\rho_t$, we build a probabilistic graphical model.
Our model assumes we have noisy information about agents, and each agent moves with some intention through the world in a way that is roughly approximated by a model.
Our model can be divided into three parts:
\begin{enumerate}
	\item Reality: This is parametrized by the true position for all time, $x_t$, and the initial velocity of the agent $v_0$.
	\item Measurements:  This is represented by our sensor readings $\hat{x}_0$ and $\hat{v}_0$ and are independent of all other variables given the true initial position and velocity, $x_o, v_0$.
	\item Motion Model: This is represented by a trajectory $\check{x}_t$ and depends on a variety of other variables which are described below.
\end{enumerate}
We now elaborate on these three components, and relate them to one another.
Many of the choices we make are motivated by a balance between model quality and computational speed (see \S \ref{sec:efficient}).

\subsection{The Variables of the Model}
The model concerns the position of an agent $x_t \in \mathbb{R}^2$ for $t \in [0,T]$ for some $T>0$.
We denote the position and velocity at time $t=0$ by $x_0$ and $v_0$ respectively.
At $t=0$, we obtain a measurement of position and velocity, denotes by $\hat{x}_0$ and $\hat{v}_0$.
Lastly, we have a variety of motion models, parametrized by a set $\mathcal{M}$ (described in the sequel).
For each model $m \in \mathcal{M}$, a trajectory $\check{x}_t$ given the initial position and velocity $x_0$ and $v_0$.
All these variables are probabilistically related to one another in a (sparse) Bayesian network, which we will describe next.

\subsection{The Sensor Model}
At time $t=0$, we obtain a noisy reading of position, $\hat{x}_0 \in \mathbb{R}^2$.
We assume that given the true position, $x_0 \in \mathbb{R}^2$, that the measurement $\hat{x}_0$ is independent of all other variables and the posterior $\Pr( \hat{x}_0 \mid x_0)$ is known.
We assume a similar measurement model for the measured initial velocity $\hat{v}_0$.

\subsection{The Agent Model}
All agents are initialized within some rectangular region $D \subset \mathbb{R}^2$.
We denote the true position of an agent by $x_t$.
We should never expect to know $x_t$ and the nature of its evolution precisely, and any model should account for its own (inevitable) imprecision.
We do this by fitting a deterministic model to the data and then smoothing the results.
Specifically, our motion model consists of a modeled trajectory $\check{x}_t$, which is probabilistically related to the true position by $x_t$ via a known and easily computable posterior, $\Pr(x_t \mid \check{x}_t)$.

Once initialized, agents come in two flavors: linear and nonlinear.
The linear agent model evolves according to the equation $\check{x}_t = x_0 + t v_0$ and so we have the posterior:
\begin{align}
	\Pr( \check{x}_t  \in A \mid x_0, v_0, lin) = \int_A \delta( \check{x}_t - x_0 - t v_0 ) d\check{x}_t.
\end{align}
for all measurable sets $A \subset \mathbb{R}^2$, where $\delta( \cdot )$ denotes the multivariate Dirac-delta distribution.
For the sake of convenience, from here on we drop the set $A$ and the integral when defining such posteriors since this equation is true for all measurable sets $A$.
We also assume the posteriors, $\Pr(x_0 \mid lin)$ and $\Pr( v_0 \mid lin, x_0)$ are known.

If the agent is of nonlinear type, then we assume the dynamics take the form:
\begin{align}n
	\frac{d \check{x}_t}{dt} = s \cdot X_k(\check{x}_t) \label{eq:ode}
\end{align}
where $X_k$ is a vector-field\footnote{A vector-field, generally speaking, is an assignment of a velocity to each position in some space.  A vector-field on $\mathbb{R}^n$ is nothing but a map from $\mathbb{R}^n \to \mathbb{R}^n$.} drawn from a finite collection $\{X_1, \dots, X_n\}$, and $s \in \mathbb{R}$.
More specifically, we assume that each $X_k$ has the property that $\| X_k(x) \| = 1$ for all $x \in D$.
This property ensures that speed is constant in time
% and it has the further advantage of being the (local) generator of solutions to an optimal navigation problem (see Appendix \ref{app:ioc}).
As we describe in \S \ref{sec:implementation}, the stationary vector-fields $X_1,\dots,X_n$ are learned from the dataset.

It is assumed that $k$ and $s$ are both constant in time, so that $\check{x}_t$ is determined from the triple $(x_0,k,s)$ by integrating \eqref{eq:ode} with the initial condition $x_0$.
This insight allows us to use the motion model to generate the posterior for $\Pr( \check{x}_t \mid x_0, k, s)$.
For each initial condition, $x_0$, we can solve \eqref{eq:ode} as an initial value problem, to obtain a point $\check{x}_t$ with initial condition $\check{x}_0 = x_0$.
This process, which takes $\check{x}_0$ and outputs $\check{x}_t$, constitutes a map which is termed the flow-map \cite[Ch 4]{MTA}, and which we denote by $\Phi_{k,s}^t$.
Explicitly, we have the posterior:
\begin{align}
	\Pr( \check{x}_t \mid x_0 , k , s) = \delta( \check{x}_t - \Phi^{t}_{k,s}( x_0) )  d\check{x}_t \label{eq:x_check | ksx}
\end{align}
where $\Phi^{t}_{k,s}$ is the flow-map of the vector field $s \,X_k$ up to time $t$.
Note that this flow-map can be evaluated for an initial condition efficiently by just integrating the vector field from that initial condition.
Note the variables $k,s$ and $x_0$ determine $v_0$.
Thus we have the posterior:
\begin{align}
	\Pr( v_0 \mid k, s, x_0) = \delta( v_0 -s X_k( x_0) ) dv_0. \label{eq:v | ksx}
\end{align}
In summary, the agent models are parametrized by the set:
\begin{align}
	\mathcal{M} = \{ lin \} \cup \left( \mathbb{R} \times \{ 1 , \dots, n \} \right) \label{eq:models}
\end{align}
and each flavor determines the type of motion we should expect from the agent model.

\subsection{The Full Model}
Concatenating the measurement model with our motion models yields the Bayesian network, where $M \in \mathcal{M}$ denotes the model of the agent:
\begin{align}
\begin{tikzpicture}[thick, var/.style={circle,draw,thin,rounded corners,shade,top color=blue!50,minimum size = 4mm}]
	\node[var] (M) {$M$};
	\node[var] (x)[right=of M] {$x_0$};
	\node[var] (v)[below=of x] {$v_0$};
	\node[var] (x_hat) [right=of x] {$\hat{x}_0$};
	\node[var] (v_hat) [right=of v] {$\hat{v}_0$};
	\node[var] (x_check_t) [left=of v] {$\check{x}_t$};
	\node[var] (x_t) [left=of x_check_t] {$x_t$};
	\draw[->] (M) to (x);
	\draw[->] (M) to (v);
	\draw[->] (M) to (x_check_t);
	\draw[->] (x) to (x_hat);
	\draw[->] (x) to (x_check_t);
	\draw[->] (x) to (v);
	\draw[->] (v) to (x_check_t);
	\draw[->] (v) to (v_hat);
	\draw[->] (x_check_t) to (x_t); 
\end{tikzpicture}.\label{eq:pgm}
\end{align}
We may use this Bayesian network to compute $\rho_t$ efficiently.
In particular
\begin{align}
	\rho_t(x_t ) &:= \Pr( x_t \mid \hat{x}_0, \hat{v}_0 ) \\
	&= \left( \sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds \right) \\
	&\quad + \Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 ). \label{eq:decomposition}
\end{align}
The final term $\Pr( x_t, lin \mid \hat{x}_0, \hat{v}_0 )$ is expressible in closed form when the posteriors $\Pr( x_0 \mid lin)$ and $\Pr( v_0 \mid lin,x_0)$ have known expressions ( e.g. Gaussians or uniform distributions). 
In this instance, the numerical computation of $\Pr(x_t, lin \mid \hat{x}_0, \hat{v}_0)$ poses a negligible burden.
Instead the primary computational burden derives from computing $\sum_{k} \int \Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 ) ds$.

\section{Efficient Probability Propagation} \label{sec:efficient}
As mentioned, many of the modeling choices are born out of a balance between accuracy and real-time computability.
One of the major modeling choices, that agents move approximately according to a small number of ODEs, is the most prominent such choice.
This section details how this modeling choice can be leveraged to compute $\rho_t(x_t)$ quickly and accurately.

To begin, rather than focusing on computing $\rho_t(x_t)$, we describe how to compute the joint probability $\Pr( x_t , \hat{x}_0, \hat{v}_0)$.
We can obtain $\rho_t(x_t)$ by normalizing $\Pr( x_t , \hat{x}_0, \hat{v}_0)$ with respect to integration over $x_t$.
We can approximate the integration over $s$ in \eqref{eq:decomposition}, with a Riemann sum.
Given a regular partition $\{ s_0, s_1, \dots, s_n \}$ of step-size $\Delta s > 0$ on the support of $\Pr(s)$, 
we can conclude that the integral term in \eqref{eq:decomposition} can be approximated by
\begin{align}
	\begin{split}
	&\sum_{k} \int \Pr( x_t, k , s , \hat{x}_0, \hat{v}_0 ) ds = \\
	 &\underbrace{\Delta s \sum_{j} \sum_{k} \Pr( x_t, k , s_j , \hat{x}_0, \hat{v}_0)}_{\text{approximation}}
	  +
	  \underbrace{\varepsilon_s}_{\text{error}}
	 \end{split} \label{eq:approximation 0}
\end{align}
where the error is bounded by $\int| \varepsilon_s | ds \leq  TV(x_t, \hat{x}_0, \hat{v}_0) \Delta s$
where $TV(x_t, \hat{x}_0, \hat{v}_0)$ is the sum, with respect to $k$, of the total variation of $\Pr( x_t, k , s  \mid \hat{x}_0, \hat{v}_0 )$ with respect to $s$ for fixed $x_t, \hat{x}_0, \hat{v}_0$.
Since this error term can be controlled, the problem of solving $\rho_t(x_t)$ is reduced to that of efficiently computing $ \Pr( x_t, k , s_j  \mid \hat{x}_0, \hat{v}_0)$
for a fixed collection of $s_j$'s.

$\hat{x}_0$ and $\hat{v}_0$ are measured and are assumed fixed for the remainder of this section.
%The only continuous variable we must discretize is $x_t$.
To begin, from \eqref{eq:pgm}  notice that:
\begin{align}
	&\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0) =\\
	&= \int \Pr( x_t, \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \\
	&= \int \Pr( x_t \mid \check{x}_t ) \Pr(\check{x}_t , \hat{x}_0, \hat{v}_0, k,s) d\check{x}_t \label{eq:convolve}
\end{align}
Observe that from the last line that $\Pr( x_t, k,s, \hat{x}_0, \hat{v}_0)$ is a convolution of the joint distribution $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$.
Assuming, for the moment, that such a convolution can be performed efficiently, we focus on computation of $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$.

Again, \eqref{eq:pgm} implies:
\begin{align}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \\
	&= \int \Pr( \check{x}_t , x_0, \hat{x}_0, v_0, \hat{v}_0, k,s) dx_0 \, dv_0 \\
	&= \int \Pr( \check{x}_t \mid  x_0, k,s, v_0) \Pr( \hat{v}_0 \mid v_0) \cdot \\
	& \hspace{10ex}  \cdot \Pr( v_0 \mid k,s,x_0) \Pr(\hat{x}_0, x_0, k, s) dx_0 \, dv_0 \nonumber \\
		&= \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right) \delta\left( v_0 - s X_k (x_0) \right) \cdot \\
		&\hspace{10ex} \cdot\Pr( \hat{v}_0 \mid v_0) \Pr(\hat{x}_0, x_0,k, s) dx_0\, dv_0, \nonumber
\end{align}
where the last equality follows from substituting \eqref{eq:x_check | ksx} and \eqref{eq:v | ksx}.
 Carrying out the integration over $v_0$ we observe:
\begin{align}
\begin{split}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right)  \cdot \\
	&\hspace{10ex} \cdot \Pr(\hat{x}_0, x_0, k, s) \Psi(\hat{v}_0 ;k,s,x_0) dx_0,
\end{split}
\label{eq:push forward}
\end{align}
 where $\Psi( \hat{v}_0 ;k,s,x_0) := \left. \Pr( \hat{v}_0 \mid v_0) \right|_{v_0 = s X_k(x_0)}$.
 Assuming the density $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0)$ is of bounded variation in the variable $x_0$ (with all other variables held fixed),
 we may approximate it as a sum of weighted Dirac-delta distributions supported on a regular grid.
 %This is merely the definition of \emph{Riemann integrable} \cite{Rudin1991}.

 To accomplish this, let $S_L(\hat{x}_0)$ denote the square of side length $L>0$ centered around $\hat{x}_0$.
 Choose  $L>0$ to be such that $\int_{S_L(\hat{x}_0)} \Pr( x_0 \mid \hat{x}_0) dx_0 = 1 - \varepsilon_{tol}$ for some error tolerance $\varepsilon_{tol}>0$.
 Then, for a given resolution $N_x \in \mathbb{N}$ define the regular grid on $S_L(\hat{x}_0)$ as $\Gamma_L( \hat{x}_0 ; N_x ) := \left\{ x_0^{i,j} \mid i,j = -N_x,\dots,N_x \right\}$, where $x_0^{i,j} = \hat{x}_0 + \frac{L}{2N_x}(i,j)$.
 The grid spacing is given by $\Delta x = ( \frac{L}{2N_x}, \frac{L}{2N_x} )$.
Approximate the smooth distribution $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ as a weighted sum of Dirac-deltas (in the variable $x_0$) supported on $\Gamma_L( \hat{x}_0;N)$:
\begin{align}
	\begin{split}
		&\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 ) = \\
		&\qquad \underbrace{\left( \sum_{i,j=-N}^{N} W(k,s,i,j,\hat{x}_0) \delta( x_0 - x_0^{i,j} ) \right)}_{\text{approximation}}
		+ \underbrace{ \varepsilon_0(x_0) }_{\text{error}}
	\end{split}
	\label{eq:approximation 1}
\end{align}
where $W(k,s,i,j,\hat{x}_0)$ is merely the evaluation of $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ at the grid point $x_0 = x_0^{i,j} \in \Gamma( \hat{x}_0; N)$.
%That is to say
%\begin{align}
%	\begin{split}
%	&W(k,s,i,j,\hat{x}_0) =\\
%	&\quad \left. \left[ \Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0; k,s,x_0 )  | \Delta x | \right] \right|_{x_0 = x_0^{i,j}} 
%	\end{split}\label{eq:constants} 
%\end{align}
For fixed $k$ and $s$, the expression $\Pr(\hat{x}_0, x_0, k, s) \Psi( \hat{v}_0 ; k, s, x_0 )$ is a density in $x_0$ and the error term in \eqref{eq:approximation 1} has a magnitude of $\| \varepsilon_0 \|_{L^1} \sim \mathcal{O}( | \Delta x |  + \varepsilon_{tol} )$ with respect to the $L^1$-norm in $x_0$.
%In particular, a feasible constant of proportionality would be the BV-norm.
%The function $W(k,s,i,j, \hat{x}_0)$, can be computed efficiently as product of the posteriors represented in our Bayesian network \eqref{eq:pgm}.
Substitution of \eqref{eq:approximation 1} into the final line of \eqref{eq:push forward} yields:
\begin{align}
	\begin{split}
	&\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s) = \\
	&\quad \sum_{i,j} W(k,s,i,j,\hat{x}_0) \delta \left( \check{x}_t - \Phi_{k,s}( x_0^{i,j}) \right) + \varepsilon_t( \check{x}_t)
	\end{split}
\end{align}
where $\varepsilon_t( \check{x}_t) = \int \delta\left( \check{x}_t - \Phi_{k,s}^{t}( x_0) \right)  \varepsilon_0(x_0) dx_0$.
The first term is computable by flowing the points of the grid, $\Gamma_L(\hat{x}_0; N_x)$, along the vector field $s X_k$.
The second term, $\varepsilon_t$, may be viewed as an error term.
In fact, this method of approximating $\Pr( \check{x}_t , \hat{x}_0, \hat{v}_0, k,s)$ as a sum of Dirac-delta distributions is adaptive, in that the error term does not grow in total mass, which is remarkable since many methods for linear evolution equations accumulate error exponentially in time \cite{Leveque1992,Gottlieb2001}:

\begin{thm} \label{thm:error}
	The error term, $\varepsilon_t \sim \mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ in the $L^1$-norm, for fixed $k,s,\hat{x}_0$, and $\hat{v}_0$.
	Moreover, $\| \varepsilon_t \|$ is constant in time.
\end{thm}
\begin{proof}
	To declutter notation, let us temporarily denote $\Phi_{k,s}^t$ by $\Phi$.
	We observe
\begin{align*}
	\| \varepsilon_t \|_{L^1} &= \int \left| \int \delta( \check{x}_t - \Phi(x_0) ) \varepsilon_0(x_0) dx_0 \right| d\check{x}_t \\
	&= \int \det( \left. D\Phi \right|_{\Phi^{-1}( \check{x}_t) }) |\varepsilon_{0}( \Phi^{-1}( \check{x}_t )) | d\check{x}_t \\
	&= \int | \varepsilon_0( u) | du = \| \varepsilon_0 \|_{L^1}
\end{align*}
As $\varepsilon_0$ is of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$ the result follows.
\end{proof}

While this allows us to compute posteriors over the output of our models, $\check{x}_t$, we ultimately care about densities over the true position.
The following corollary of Theorem \ref{thm:error} addresses this:
\begin{cor} \label{cor:error}
	The density
	\begin{align}
		\sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^\alpha) } \label{eq:approximation 2}
	\end{align}
	is an approximation of $\Pr( x_t, k, s, \hat{x}_0, \hat{v}_0)$ with a constant in time error bound of magnitude $\mathcal{O}( |\Delta x| + \varepsilon_{tol} )$.
\end{cor}
\begin{proof}
	By \eqref{eq:convolve}
	\begin{align*}
		\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) = \int \Pr( x_t \mid \check{x}_t ) \Pr( \check{x}_t, k,s,\hat{x}_0, \hat{v}_0) d\check{x}_t
	\end{align*}
	Substitution of \eqref{eq:approximation 1} yields
	\begin{align*}
		\begin{split}
		&\Pr(x_t, k, s, \hat{x}_0, \hat{v}_0) \nonumber \\
		& = \sum_{i,j} W(k,s,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,s}^t( x_0^{i,j}) } + \tilde{\varepsilon}_t(x_t)
		\end{split}
	\end{align*}
	where the error term is
	\begin{align}
		\tilde{\varepsilon}_t(x_t) = \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t
	\end{align}
	and $\varepsilon_t$ is the error term of \eqref{eq:approximation 1}.
	We see that the $L^1$-norm of $\tilde{\varepsilon}_t$ is 
	\begin{align}
		\| \tilde{\varepsilon}_t \|_{L^1} &= \int \left| \int \Pr( x_t \mid \check{x}_t) \varepsilon_t( \check{x}_t) d \check{x}_t \right| dx_t \\
			&\leq \int \Pr( x_t \mid \check{x}_t ) | \varepsilon_t |( \check{x}_t) d\check{x}_t \, dx_t
	\end{align}
	Implementing the integration over $x_t$ first yields:
	\begin{align}
		\| \tilde{\varepsilon}_t \|_{L^1} \leq \int | \varepsilon_t |( \check{x}_t) d\check{x}_t =: \| \varepsilon_t \|_{L^1}
	\end{align}
	which is $\mathcal{O}( | \Delta x | + \varepsilon_{tol} )$ by Theorem \ref{thm:error}.
\end{proof}

Corollary \ref{cor:error} justifies using \eqref{eq:approximation 2} as an approximation of $\Pr( x_t, k,s,\hat{x}_0, \hat{v}_0)$.
In particular, this reduces the problem of computing $\rho_t(x_t)$ to the problem of computing the weights $W(k,s,i,j,\hat{x}_0)$ and the points $\Phi_{k,s}^t(x_0^{i,j})$ for all $k,s$ and points $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
We can reduce this burden further by exploiting the following symmetry:
\begin{thm} \label{thm:symmetry}
	$\Phi_{k,s}^t = \Phi_{k,1}^{st}$.
\end{thm}
\begin{proof}
	Say $x(t)$ satisfies the ordinary differential equation $x'(t) = sX_k(x(t))$ with the initial condition $x_0$.
	In other words, $x(t) = \Phi_{k,s}^{t}(x_0)$.
	Taking a derivative of $x(t/s)$, we see $\frac{d}{dt} (x(t/s)) = x'(t/s) /s = X_k(x(st))$.
	Therefore $x(t/s) = \Phi_{k,1}^{t}( x_0)$.
	Substitution of $t$ with $\tau = t/s$ yields $x(\tau) = \Phi_{k,1}^{s \tau} (x_0)$.
	As $x(\tau) = \Phi_{k,s}^{\tau}(x_0)$ as well, the result follows.
\end{proof}
This, allows us to compute $\Phi_{k,s}^t( x_0^{\alpha})$ using computations of $\Phi_{k,1}^t(x_0^{\alpha})$.
This yields the following theorem
\begin{thm} \label{thm:main}
	Let $\{s_1,\dots,s_n\}$ be a regular partition on the support of $\Pr(s)$.
	Then the density
	\begin{align}
		\begin{split}
		&\Delta s \sum_{i,j,k,\ell} W(k,s_\ell,i,j,\hat{x}_0) \left. \Pr( x_t \mid \check{x}_t ) \right|_{ \check{x}_t = \Phi_{k,1}^{s_\ell t}( x_0^{i,j}) } \\
		&+\Pr( x_t , lin, \hat{x}_0, \hat{v}_0 )
		\end{split}
		\label{eq:approximation 3}
	\end{align}
	approximates $\Pr( x_t, \hat{x}_0, \hat{v}_0)$ with an error of size $\mathcal{O}( \Delta s + \Delta x + \varepsilon_{\rm tol})$.
\end{thm}
\begin{proof}
	Substitute Theorem \ref{thm:symmetry} into Corollary \ref{cor:error}, to replace $\Phi_{k,s}^t$ with $\Phi_{k,1}^{st}$.
	This gives us an error term of size $\Delta x$, if we compute the integral over $s$ exactly.
	Using \eqref{eq:approximation 0}, we can compute the integral over $s$ approximately, with an error of magnitude $\Delta s$.
\end{proof}

This is a powerful result, since if $\Pr(s)$ has a support contained within the domain $[-\bar{s}, \bar{s}]$, then to solve for $\Pr( x_t, \hat{x}_0, \hat{v}_0)$  (and thus $\rho_t(x_t)$) at times $0, \Delta t, \dots, N_t \Delta t$, we only need to compute $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ for $\ell = -N_t,\dots,N_t$ once.
This is because, once $\Phi_{k,1}^{\ell \Delta t \bar{s}}$ is computed for $\ell = -N_t, \dots ,N_t$ we can use the partition of $[-\bar{s}, \bar{s}]$ given by $\{ \bar{s} \frac{m}{\ell} \}_{m=-\ell}^{\ell}$ to compute a $\Delta s = \bar{s} / \ell$ accurate approximation for each $\ell$, courtesy of Theorem \ref{thm:main}.
%Admittedly, this means our approximations might be imprecise over very short times, but more accurate over longer time-horizons.
The procedure to compute   $\rho_t(x_t)$ is summarized in Algorithm \ref{alg:1}.

\begin{algorithm}[t]
		  \caption{Algorithm to Compute  $\rho_t$.}
		  \label{alg:1}
		  \begin{algorithmic}[1]
		    \Require $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.
		    \ForEach {$m \in \{ -\ell,\dots,\ell \}$}
		    \State Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$.
		    \EndFor  
		    \State Construct a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$ via Theorem \ref{thm:main} \label{alg1:step2}.
		    \State Normalize with respect to $\check{x}_t$ to obtain an approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$.
		    \State Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$
		  \end{algorithmic}
\end{algorithm}

%\begin{algorithm}[t]
%		    \KwData{ $N_t \in \mathbb{N},  \{ X_k \}_{k=1}^{n}, \hat{x}_0, \hat{v}_0, \Gamma_L( \hat{x}_0, N_x)$, and the posteriors of \eqref{eq:pgm}.}
%		    \For{$\ell \in \{ -N_t, \dots, N_t\}$}{
%		    	Compute $\Phi_{k,1}^{\bar{s} \ell \Delta t}( x_0^{i,j} )$ for all $x_0^{i,j} \in \Gamma_L( \hat{x}_0,N)$ \;
%			\For{$m \in \{ -\ell,\dots,\ell \}$}{
%				Compute $W(k,s,i,j,\hat{x}_0)$ for $s = \bar{s} m / \ell$ \;
%			}
%			Use Theorem \ref{thm:main} to obtain a $\mathcal{O}( \Delta x + \bar{s}/\ell + \varepsilon_{\rm tol})$ approximation of $\Pr(\check{x}_t, \hat{x}_0, \hat{v}_0)$ at $t=\ell \Delta t$\;
%			Normalize with respect to $\check{x}_t$ to obtain an equally accurate approximation of $\Pr(\check{x}_t \mid \hat{x}_0, \hat{v}_0)$ \;
%			Convolve with respect to $\Pr( x_t \mid \check{x}_t)$ to obtain $\rho_t(x_t)$ at $t= \ell \cdot \Delta t$.
%		    }
%		    \KwResult{An approximation of $\rho_t(x_t)$ at $t= 0,\Delta t, 2 \Delta t, \dots, N_t \Delta t$.}
%		    \caption{Our algorithm for computing $\rho_t$}
%		    		    \label{alg:main}
%\end{algorithm}

%Steps \ref{alg1:step1} and \ref{alg1:step2} are where the bulk of the work occurs.
%Fortunately, these two steps are embarrassingly parallel and yield tolerable computational complexity for real-time computation.
For fixed $k$ $i$ and $j$, the computation of $\Phi_{k,1}^t(x_0^{i,j})$ at $t = -N_t\bar{s}, (1-N_t)\bar{s} ,\dots, N_t \bar{s}$ takes $\mathcal{O}(N_t)$ time using an explicit finite difference scheme and can be done in parallel for each $k \in \{1,\dots, n\}$ and $x_0^{i,j} \in \Gamma_L( \hat{x}_0; N_x)$.
Similarly, computing $W(k,s,i,j,\hat{x}_0)$ constitutes a series of parallel function evaluations over tuples $(k,s,i,j)$, where the continuous variable $s$ is only required at finitely many places in Algorithm \ref{alg:1}.
If the posteriors represented by the arrows in \eqref{eq:pgm} are efficiently computably then the computation of $W(k,s,i,j, \hat{x}_0)$ is equally efficient.
%We summarize this in the following theorem:
%\begin{thm}
%	Suppose we have an accurate scheme for integrating ODEs in a computation time proportional to the integration time-interval\footnote{For example, 4th order explicit Runge-Kutta}.  
%	Algorithm \ref{alg:main} yields an $\mathcal{O}( \Delta x + \bar{s} / \ell + \epsilon_{\rm tol})$ accurate approximation of $\rho_t$ in $\mathcal{O}( N_t n N_x^2 / N_p )$ computation time given $N_p$ processing units and $n$ vector fields.
%\end{thm}

\section{Implementation and Experimental results} \label{sec:implementation}
  Given this model, we describe a specific implementation and the process of fitting the model to a dataset.
  In particular, we must learn the vector fields $\{X_1, \dots, X_n\}$, the posteriors $\Pr( x_0 \mid M)$ and the priors $\Pr(M)$ for $M \in \mathcal{M}$ from the data.
  For the purposes of demonstration, we use the Stanford Drone Dataset \cite{Robicquet2016}.
  More generally, we assume that for a fixed scene we have a database of previously observed trajectories $\{ \hat{x}^1, \dots, \hat{x}^N\}$.
  From this data we tune the parameters of the model ($\{X_1, \dots, X_n\}$, $\Pr( x_0 \mid M)$ and $\Pr(M)$)  appropriately.

  \subsection{Learning the Vector Fields}
  Before we begin to learn vector-fields, we must learn the number of possible vector-fields.
  To do this we use a clustering algorithm on the trajectories observed in a scene to categorize them into groups.
  %In principal, any clustering algorithm could accomplish this.
  In particular, we use the start and end point for each trajectory to obtain a point in $\mathbb{R}^4$.
  We then cluster in $\mathbb{R}^4$ using Affinity propagation \cite{FreyDueck2007}.
  This clustering of the end-points induces a clustering of the trajectories.
  Suppose we obtain clusters $S_1, \dots, S_n$ consisting of trajectories from our data set, as well as a set of unclassified trajectories, $S_0$.

  For each set $S_k$ we learn a vector-field that is approximately compatible with that set.
  %We'd like to only consider vector-fields who's vectors have unit magnitude because we observe many of the trajectories in the data appear to have a roughly constant speed.
  Since most trajectories appearing in the dataset have roughly constant speed, we chose a vector-field that has unit magnitude. 
  That is, we assume the vector-field takes the form $X_k(x) = \left( \cos( \Theta_k(x) ) , \sin(\Theta_k(x)) \right)$ for some scalar function $\Theta_k(x)$.
  Learning the vector-fields then boils down to learning the scalar function $\Theta_k$.
  We assume $\Theta_k$ takes the form
  \begin{align*}
  	\Theta_k(x) = \sum_{\alpha} \theta_{k,\alpha} L_{\alpha}(x),
  \end{align*}
  for some collection of coefficients, $\theta_{k,\alpha}$, and a fixed collection of basis functions, $L_{\alpha}$.
  In our case, we choose $L_{\alpha}$ to be a set of low degree Legendre polynomials.
 $\Theta_k$ can be learned by computing the velocities observed in the cluster, $S_k$.
  These velocities may be obtained by a low order finite difference formula.
  Upon normalizing the velocities, we obtain a unit-length velocity vectors, $v_{i,k}$, anchored at each point, $x_{i,k}$, of $S_k$.
  We learn $\Theta_k$ by defining the cost-function:
  \begin{align*}
  	C[ \Theta_k] = \sum_i \langle v_{i,k} , ( \cos(\Theta_k( x_{i,k}) , \sin( \Theta_k( x_{i,k} ) ) \rangle
  \end{align*}
  which penalizes $\Theta_k$ for producing a misalignment with the observed velocities at the observed points of $S_k$.
  When $\Theta_{k}$ includes high order polynomials (e.g. beyond 5th order), we also include a regularization term to bias the cost towards smoother outputs.
  Using the $H^1$-norm times a fixed scalar suffices as a regularization term.

  \subsection{Learning $\Pr( x_0 \mid M)$ and $\Pr(M)$}

  We begin by considering the nonlinear models first.
%  That is to say, we are concerned with $M = (k,s)$ for an integer $k$ and scalar $s$.
%  We'd like to compute $\Pr( x_0 \mid k,s)$ for each cluster $k=1,\dots,n$ and each speed $s \in \mathbb{R}$.
  We make the modeling assumption that $x_0$ is independent of $s$ given $k$, i.e. $\Pr( x_0 \mid k,s) = \Pr(x_0 \mid k)$.
  Additionally, we assume that $s$ and $k$ are independent.
  This means that we only need to learn $\Pr( x_0 \mid k)$, $\Pr(k)$, and $\Pr(s)$.

  We let $\Pr(k) = (n+1)^{-1}$ and $\Pr(s) \sim \mathcal{U}( [-s_{\max}, s_{\max} ] )$  where $s_{\max}>0$ is the largest observed speed in the dataset.  
  This implies that $\Pr(lin) = (n+1)^{-1}$ as well.
  Other reasonable choices (such as $\Pr(k) \propto | S_k|$) could work, but we have chosen to be conservative in this paper.

%  To learn $\Pr( x_0 \mid k)$, we assume that its logarithm may be approximated by a low dimensional subspace of functions on $\mathbb{R}^2$.
%  That is to say, 
  For each $k$ we assume $\Pr( x_0 \mid k) = \frac{1}{Z_k} \exp( - V_k(x_0) )$ and $V_k$ is a function whose constant term is $0$ and is given by:
  \begin{align*}
  	V_k(x_0; \mathbf{c} ) := \sum_{|\alpha|< d} c_{\alpha} L_{\alpha}( x_0)
  \end{align*}
  for a collection of basis functions, $L_{\alpha}$ and coefficients $\mathbf{c} = \{ c_{\alpha} \}_{|\alpha| < d}$.
  We chose our basis functions to be the collection of tensor products from the first six Legendre polynomials, normalized to the size of the domain.
  Then, one may fit the coefficients $c_{\alpha}$ to the data by using a log-likelihood criterion.
  The resulting (convex) optimization problem takes the form:
  \begin{align*}
  	\mathbf{c}^* = \inf_{ |\mathbf{c}| } \sum_{x \in S_k} V_k( x_0; \mathbf{c})
  \end{align*}
  Where the norm on $\mathbf{c}$ is a sup-norm.
  We can also bias this optimization towards more regular functions by adding a penalty to the cost function.
    Finally, we let $\Pr( x_0 \mid lin) \sim \mathcal{U}(D)$.

  \subsection{Learning the Measurement Model}
  We assume a Gaussian noise model.  
  That is
  \begin{align*}
  	\Pr( \hat{x}_0 \mid x_0 ) \sim \mathcal{N}( x_0 , \sigma_x) \quad, \quad \Pr( \hat{v}_0 \mid v_0 ) \sim \mathcal{N}( v_0, \sigma_v).
  \end{align*}
  Therefore, our model is parametrized by the standard deviations $\sigma_x$ and $\sigma_v$.
  We assume that the true trajectory of an agent is smooth compared to the noisy output of our measurement device.
  This justifies smoothing the trajectories, and using the difference between the smoothed signals and the raw data to learn the variance $\sigma_x$.
  To obtain the results in this paper we have used a moving average of four time steps (this is $0.13$ seconds in realtime).
  We set $\sigma_v = 2 \sigma_x / \Delta t$ where $\Delta t > 0$ is the time-step size.  
  This choice is justified from the our use of finite differences to estimate velocity.
  In particular, if velocity is approximated via finite differencing as $v(t) \approx (x(t+h) - x(t))\,\Delta t^{-1} + \mathcal{O}(h)$ and the measurements are corrupted by Gaussian noise, then the measurement $\hat{v}(t)$ is related to $v(t)$ by Gaussian noise with roughly the same standard deviation as $(x(t+h) - x(t))\,\Delta t^{-1}$.

  \subsection{Learning the Noise Model}
  Finally, we assume that the true position is related to the model by Gaussian noise with a growing variance.
  In particular, we assume $\Pr( x_t \mid \check{x}_t) \sim \mathcal{N}( \check{x}_t , \kappa t)$ for some constant $\kappa \geq 0$.
  The parameter, $\kappa$, must be learned.
 % We've taken an adhoc approach in this paper, although more intelligent and theoretically sound alternatives likely exist.
  For each curve in $S_k$ we create a synthetic curve using the initial position and speed and integrating the corresponding vector-field, $s\, X_k$.
  So for each curve, $x_i(t)$, of $S_k$, we have a synthesized curve $x_{i,synth}(t)$ as well.
  We then measure the standard deviation of $(x_i(t) - x_{i,synth}(t)) / t$ over $i$ and at few time, $t \in \{ 0, 100, 200 \}$ in order to obtain $\kappa$.

 \subsection{Evaluating Performance}
In this section we establish our methods for evaluating the quality of our predictions, and comparing them against the model from [8] and a random walk.
We implemented our model as well as the evaluation code in Python 2.6. Our implementation is available online \footnote{https://github.com/OkayHughes/iros2017\_pedestrian\_forecasting}.

\begin{figure}
	\centering
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Our Algorithm}}   
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/ours_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Kitani et al.}} 
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/kit_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Random Walk} }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=70.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=250.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/comparison/rand_death_1_2_t=370.png}
	\end{minipage}

	\vspace{0.2cm}
	\hspace{.5cm}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 2.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 8.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 12.33s$
	\end{minipage}

	\caption{We behave qualitatively and quantitatively better than kitani in this example. Our algorithm starts out with a dense distribution, which spreads with time, whereas Kitani et al.'s algorithm is very confident in its mistake.}

	\label{fig:death-1-2}
\end{figure}

\begin{figure}
	\centering
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Our Algorithm}   }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/ours_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Kitani et al.}  }
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/kit_1_2_t=370.png}
	\end{minipage}

	\vspace{0.1cm}
	\begin{minipage}[c]{0.5cm}
		\rotatebox{90}{\small{Random Walk} }  
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=50.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=130.png}
	\end{minipage}
	\begin{minipage}[c]{0.3\linewidth}
		\includegraphics[width=\linewidth]{./figures/bookstore/rand_1_2_t=370.png}
	\end{minipage}

	\vspace{0.2cm}
	\hspace{.5cm}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 1.66s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 4.33s$
	\end{minipage}
	\begin{minipage}{0.3\linewidth}
		\centering
		$t = 12.33s$
	\end{minipage}

	\caption{We predict this trajectory very well, and while Kitani et al. captures the same behaviour, however their motion model doesn't adequately keep up with the ground truth of the pedestrian.}

	\label{fig:bookstore-1-2}
\end{figure}

Our data came from the Stanford Drone Dataset in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles). 
Our test data used the bicycle annotations. 
We did a 2-fold cross validation on 4 different scenes of varying complexity. 
Our analysis used the same partitions of the trajectories for all three predictors.

Our data came from the Stanford Drone Dataset, in the form of hand annotated bounding boxes around agents (e.g. cars, pedestrians, bicycles). Our datasets used the bicycle data. We did a 2-fold cross validation on 4 different scenes of varying complexity. Our analysis used the same partitions of the trajectories for all three predictors.

The algorithm provided by Kitani et al. was given additional information when making predictions that neither ours nor the random walk model were provided, namely the endpoint of the test trajectory. Without these data the implementation of the predictor provided by the authors devolves into a random walk.

The output distributions of the three algorithms were compared using their integrals over the cells of a suitably fine regular grid over our domain. These integrals are used to visualize the distributions in Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2}.
In keeping with our measurement model, the ground truth compared against was a grid that was identically zero except on a bounding box around the measurement at time $t$. The dimensions of the bounding box were the average bounding box dimensions in the training data set. Because our predictions all share the same scale, we were able to amalgamate all of the prediction and truth values for all of the simulated agents at a given time step and generate ROC curves. Figure \ref{fig:auc_vs_time} shows the Area Under the Curve of these ROC curves versus time.
\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{./figures/the_results.eps}
	\caption{We see that at all times we outperform Random Walk and Kitani et al.'s algorithm. The initial dip in Kitani et al.'s algorithm is due to their initial distribution being very sharp.}
	\label{fig:auc_vs_time}
\end{figure}

Disregarding small time scales where any algorithm will be reasonably accurate, our predictor behaves much better than any of the others at moderately large $t$, where we capture the  behavior of the agent without suffering the misplaced confidence observed in Kitani et al.'s predictions. Their encoding of velocity results in the prediction lagging behind the actual observed motion. Even given significantly better information, Kitani et al. generates lower quality predictions than our algorithm at all time scales. 
\begin{table}
	\begin{center}
		\caption{Runtimes}
		\label{tab:time}
		\renewcommand{\arraystretch}{1.5}% Spread rows out...
		\begin{tabular}{||c | c c  c ||} 
			\hline
			& Our Predictor & Random Walk & Kitani et al. \\ [0.5ex] 
			\hline 
			$\frac{\mathrm{time}}{\mathrm{frame}}$ & 0.0169s & 2.1E-7s & 0.0614s \\
			\hline

		\end{tabular}

	\end{center}
\end{table}

The run time per frame for each algorithm was generated using total run time for 400 frames, averaged across several agents and scenes, divided by the number of frames. This is shown in Table~\ref{tab:time}. Our algorithm implementation leveraged its embarassing parallelism using a relatively naive parallelization scheme, which split the computation of frames among 18 cores. It's highly probable that our run times would significantly decrease using an implementation of our algorithm in C that uses a more clever scheme. Kitani et al. was timed with minimal modification to the source code provided by the authors.

Figure \ref{fig:auc_vs_time} shows that at all time scales our algorithm has better predictive capability than Kitani et. al's. We also run over 3 times faster than their algorithm, at over 60FPS.  Figures \ref{fig:death-1-2} and \ref{fig:bookstore-1-2} show the qualitative superiority of our predictions on several scenes.

\section{Conclusion} 
\label{sec:conclusion}

\Ram{Matt can you please take a crack at this ....}
Avenues of improvement:
update predictions when given additional measurements

The conclusion goes here.

\section*{Acknowledgments}
Kris Kitani was critical in aiding the comparison to his algorithm.

%\appendix
%\subsection{Inverse Optimal Control} \label{app:ioc}
%One of the valuable aspect of the motion model of \cite{Kitani2012} was that the trajectories were solutions of optimal control problems.
%This is a desirable property for a motion model of pedestrians/cyclists/vehicles in that it seems reasonable to assume that we all move through the world with an intent to get somewhere.
%This framework was dubbed \emph{inverse optimal control} (IOC).
%Here we illustrate how solutions of \eqref{eq:ode} can fit within the IOC framework.
%
%\begin{thm}
%	Let $X$ be a vector-field such that $\| X(x) \|_2 = 1$ for all $x \in \mathbb{R}^2$.
%Then, for each initial condition $x_0 \in \mathbb{R}^2$, there exists a neighborhood $U$, a Riemannian metric, $g$, and a cost function, $f$, 
%such that solutions of \eqref{eq:ode} are also solutions of the optimal control problem
%\begin{align}
%	p^* = \inf_{ \| v_t \|_g = s } \int_0^T f( x_t) dt \label{eq:IOC}
%\end{align}
%where $v_t = \frac{d x_t}{dt}$, $\| \cdot \|_g$ is the norm with respect to $g$ and the infimum is taken over all differentiable curves in $U$ which originate from $x_0$.
%\end{thm}
%
%\begin{proof}
%	By rescaling time, we may let $s$ be any positive real number.
%	Therefore, without loss of generality, we will seek an $f$ and $g$ such that $s=1$.
%	In this case a solution of \eqref{eq:IOC} is generated by the vector-field $X = \frac{\nabla f}{ \| \nabla f\|}$ where $\nabla$ is the Riemannian gradient with respect to some metric (possibly non-Euclidean).
%	Therefore, our goal is to illustrate that there exists a metric $g$ and a function $f$ such that the given $X$ is given by $\frac{\nabla f}{ \| \nabla f\|}$.
%
%	Because $\| X \| = 1$ globally, it has no fixed points.
%	From the flow-box theorem \cite[Theorem 4.1.14]{MTA} we can assert that for any open set, $U \subset \mathbb{R}^2$ there exists a local diffeomorphism which transforms $X_k$ into a flat vector-field.
%	Mathematically, this asserts the existence of a map $\Phi : U \to V \subset \mathbb{R}^2$ such that the push-forward of $X$, which we will denote by $\tilde{X}$, and given by
%	\begin{align*}
%		\tilde{X}(\tilde{x}) = \Phi_* X(\tilde{x}) := \left. D\Phi \right|_{\Phi^{-1}(\tilde{x}) } \cdot X(  \Phi^{-1}(\tilde{x}) ),
%	\end{align*}
%	is just the flat vector field $(1,0)$ for all $x \in U$.
%	We could view the coordinate function $\tilde{f}(x) = x^0$ as a cost function on $V$,
%	and we may set $g_V$ to be equal to the standard flat metric on $V$ inherited from $\mathbb{R}^2$.
%	In this case we observe that $\tilde{X}$ generates solutions to the optimization problem
%	\begin{align*}
%		\tilde{p}^* = \inf_{ \| \tilde{v}_t \|_{g_V} \leq 1} = \int_0^T \tilde{f}( \tilde{x}_t ) dt
%	\end{align*}
%	By a change of variables, it follows that the original vector field, $X$, then solves the optimization problem \eqref{eq:IOC}
%	where $g = \Phi^* g_V$ is the pull-back metric, and $f = \Phi^* \tilde{f}$ is the pull-back of $\tilde{f}$.
%\end{proof}

%% Use plainnat to work nicely with natbib.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}